---
title: "Wk5Exercise"
author: "Matt Ming"
date: "2/21/2022"
output: html_document
---

# Set up

Loading in the tidyverse packages to enable the later functions
```{r,message=FALSE,warning=FALSE,error=FALSE}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(mosaic)
```

# Challenge 1

First, let's load the IMDb dataset
```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2022-datasets/master/IMDB-movies.csv"
data <- read_csv(f,col_names = TRUE)
head(data)
```

Now, filter the data by only those movies with runtime between 1 - 3 hours
and made between the years 1920 - 1979.  We should be left with a dataframe with
5651 observations
```{r}
data <- data %>% 
  filter(runtimeMinutes %in% 60:180 & startYear %in% 1920:1979) %>%
  mutate(decade = case_when(startYear %in% 1920:1929 ~ "20s",
                            startYear %in% 1930:1939 ~ "30s",
                            startYear %in% 1940:1949 ~ "40s",
                            startYear %in% 1950:1959 ~ "50s",
                            startYear %in% 1960:1969 ~ "60s",
                            startYear %in% 1970:1979 ~ "70s"))
```

Now, using ggplot to get a histogram of the distribution of runtimeMinutes for
each decade. (Hint: use facet_wrap())
```{r}
p <- ggplot(data,aes(runtimeMinutes)) + facet_wrap(~decade,ncol = 3) + 
  geom_histogram(binwidth = 5)
print(p)
```

Getting the mean and standard deviation for movie runtimes in minutes per decade
then save the calculations in a new dataframe called **results**
```{r}
results <- data %>% select(runtimeMinutes,decade) %>% group_by(decade) %>%
  summarize(runtimes_means = mean(runtimeMinutes),
            runtimes_sds = sd(runtimeMinutes))
```

Drawing samples of 100 movies from each decade and calculating the single-sample
mean and sd in runtimeMinutes per decade.  We can then use the single-sample
sd and sample size to estimate the standard error by the formula  
$SE\approx\sigma_x/\sqrt{n}$ where $\sigma_x$ = sd of the sample, and $n$ = 
sample size.  We then want to compare these estimates to SE estimates using the 
population mean and sd (again using sample size of 100)
```{r}
results <- results %>% mutate(sample_mean = rep(0,6),
                              sample_sd = rep(0,6),
                              runtimes_se = rep(0,6),
                              sample_se = rep(0,6))
for(i in seq(20,70,by=10)){
  decstr <- paste0(i,"s")
  samplei <- data %>% filter(decade == decstr) %>% sample_n(size=100)
  results[(i/10)-1,] <- results %>% filter(decade == decstr) %>%
    mutate(sample_mean = mean(samplei$runtimeMinutes),
                     sample_sd = sd(samplei$runtimeMinutes)) %>%
    mutate(runtimes_se = runtimes_sds/10,
           sample_se = sample_sd/10)
}
results <- results %>% select(decade,runtimes_means,sample_mean,
                              runtimes_sds,sample_sd,
                              runtimes_se,sample_se)
head(results)

```

Now, we're going to generate a sampling distribution by sampling 100 movies from
each decade 1000 times, and getting the mean for each of those 1000 samples and
plot them.  We then want to compare the standard error found from this sampling
with the SE calculated using the sample sd and population sd calculated earlier.
```{r}
for(i in seq(20,70,by = 10)){
  decstr <- paste0(i,"s")
  decsamps <- data.frame(means = rep(0,1000),sds = rep(0,1000))
  dec_pop <- data %>% filter(decade == decstr) %>% select(runtimeMinutes)
  decsamps$means <- (do(1000) * mean(sample(dec_pop$runtimeMinutes,100)))$mean
  decsamps$sds <- (do(1000) * sd(sample(dec_pop$runtimeMinutes,100)))$sd
  
  sampdist_mean <- mean(decsamps$means)
  sampdist_sd <- sd(decsamps$sds)
  print(histogram(decsamps$means,
            main = paste("Histogram of sampling distribution for movies from the",decstr),
            xlab = "Mean of sampled runtime"))
  
  print("Comparing estimates of SE")
  print(paste("Decade =",decstr))
  print(paste("Sampling Dist SE =",sampdist_sd))
  print(paste("Sample SE        =",results$sample_se[which(results$decade == decstr)]))
  print(paste("Population SE    =",results$runtimes_se[which(results$decade == decstr)]))
  cat("\n")
}
```

Visually inspecting the histograms produced, we can see that the sampling
distribution in every decade is approximately normally distributed.

# Challenge 2

We'll now switch to a second dataset, which will be loaded in here:
```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2022-datasets/main/zombies.csv"
data2 <- read_csv(f,col_names = TRUE)
head(data2)
```

First, we're going to calculate the population-level mean and standard deviation
for each quantitative field, those fields being Height, Weight, Age, Zombies
Killed, and Years of Education.  Note that because we're dealing with
*population*-level statistics here, we'll want to manually calculated the sd as
$\sigma=\sqrt{SS/N}$ where $SS$ is the sum of squares, or the deviation of each
data point from the mean.
```{r}
pop_stats <- data.frame(fields = c("height","weight","age","zombies_killed",
                                   "years_of_education"),
                        means = rep(0,5),SDs = rep(0,5))
for(i in pop_stats$fields){
  pop_stats[which(pop_stats$fields == i),"means"] <- mean(data2[[i]])
  pop_stats[which(pop_stats$fields == i),"SDs"] <- 
    sqrt(sum((data2[[i]] - mean(data2[[i]]))^2)/length(data2[[i]]))
}
head(pop_stats)

```

Now, to generate some plots using the data.  First, we make boxplots for each
of the above variables by gender.
```{r}
for(i in pop_stats$fields){
  p <- ggplot(data2,aes_string(x = "gender",y = i)) + geom_boxplot() +
    ggtitle(paste(i,"by Gender"))
  print(p)
}
```

Now, making a scatterplot of the relationship between height and age, and weight
and age.  Also, coloring each point by gender
```{r}
p1 <- ggplot(data2,aes(x = age,y = height,color = gender)) + geom_point() +
  ggtitle("Height vs Age")
p2 <- ggplot(data2,aes(x = age,y = weight,color = gender)) + geom_point() +
  ggtitle("Weight vs Age")
print(p1)
print(p2)
```

When we generate these plots, we notice a few trends.  First, we can see a very
strong positive correlation between age and height, meaning that across the 
population, in general older age means taller height.  Likewise, there is a very
slight positive correlation between weight and age but it is not nearly as
strongly positive and the clustering of points around the best-fit line is not
as tight.  We also see that height is very correlated with gender, as Males
were, on average, significantly taller (~5 inches) than Females.  Again, there
was also a slight correlation between weight and gender, with Males being
heavier on average than Females, but it is difficult to see this difference from
the scatter plot as the variance in weight given a certain age is much greater
for both genders.  The difference in average weight between Males and Females is
much more noticeable when comparing the boxplot of weight vs. gender.

We now want to examine the normality of each of the variables above.  To do this
we are going to first plot a histogram of the data and see if it looks
approximately normal, then we'll back up our findings using QQ plots.
```{r}
for(i in pop_stats$fields){
  hist(data2[[i]],main = paste("Histogram of",i),xlab = i,breaks = 20)
  qqnorm(data2[[i]],
         main = paste("Q-Q plot of",i,"folowing a normal distribution"),
         xlab = i)
  qqline(data2[[i]],col = "red")
}
```

Visually inspecting these plots, it seems that height, weight, and age follow
a normal distribution; this makes sense as each of these variables is 
continuous.  However, the number of zombies killed and years of education
variables do not follow a normal distribution because they are not continuous
but are rather discrete variables (i.e., one cannot kill 2.7 zombies).  Thus,
it appears that these variables are drawn from a Poisson distribution (each
with some unknown parameter $\lambda$)

Now, we'll sample 50 survivors and calculating the mean and standard deviation,
as well as an estimate of standard error from the sample SD for each variable.
Again, we will use our knowledge of $SE\approx\sigma_x/\sqrt{n}$ to calculate
SE at this step.  We also want to calculate a 95% confidence interval for the
single sample, which we can do by using the SE and knowing that 95% of a normal
distribution falls between $\mu\pm1.96*SE$ where $\mu$ = sample mean; note that
here we assume that the sampling distribution of means for each variable follows 
a normal distribution.

```{r}
survivor_samp <- sample_n(data2,50)
stats_tab <- data.frame(fields = c("height","weight","age","zombies_killed",
                                   "years_of_education"),
                        means = rep(0,5),SDs = rep(0,5),SEs = rep(0,5),
                        lower_ci = rep(0,5),upper_ci = rep(0,5))
for(i in 1:5){
  stats_tab[i,"means"] <- mean(survivor_samp[[stats_tab$fields[i]]])
  stats_tab[i,"SDs"] <- sd(survivor_samp[[stats_tab$fields[i]]])
  stats_tab[i,"SEs"] <- stats_tab[i,"SDs"]/sqrt(50)
  stats_tab[i,"lower_ci"] <- stats_tab[i,"means"] + 
    stats_tab[i,"SEs"]*qnorm(0.025)
  stats_tab[i,"upper_ci"] <- stats_tab[i,"means"] + 
    stats_tab[i,"SEs"]*qnorm(0.975)
}
print(stats_tab)
```

We'll now perform this same sampling 200 times to develop a sampling 
distribution for each variable, and use these sampling distributions to 
calculate SE. We will also use the SEs calculated to estimate 95% confidence
intervals for the mean of each variable.
```{r}
for(i in 1:5){
  print(paste("Variable =",stats_tab$fields[i]))
  means_i <- stats_tab$means[i]
  means_i <- c(means_i,(do(199) * 
                          mean(sample_n(data2,50)[[stats_tab$fields[i]]]))$mean)
  hist(means_i,
       main = paste("Histogram of estimates of the mean for",stats_tab$fields[i]),
       xlab = stats_tab$fields[i],breaks = 20)
  qqnorm(means_i,
         main = paste("QQ Plot of estimates of the mean for",stats_tab$fields[i]))
  qqline(means_i,col = "red")
  print(paste("Mean Sample Dist mean  =",mean(means_i)))
  print(paste("Mean Sample Dist SD    =",sd(means_i)))
  print(paste("Previous SE            =",stats_tab$SEs[i]))
  print(paste("Samp dist 95% Lower CI =",quantile(means_i,0.025)))
  print(paste("Previous Lower CI      =",stats_tab$lower_ci[i]))
  print(paste("Samp dist 95% Lower CI =",quantile(means_i,0.975)))
  print(paste("Previous Upper CI      =",stats_tab$upper_ci[i]))
  cat("\n")
}
```

Based on the QQ plots we see that all of the sampling distributions are
approximately normal with some slight deviations simply due to sample size; even
variables which themselves were not normally distributed have sampling
distributions which are normal.  The histograms do not visually appear to 
conform very closely to a normal distribution, but they still have an 
approximately normal shape with some expected deviation due to random sampling.

The standard errors we observed from the 200 samples seem to be lower than the
SEs calculated using the mean and standard deviation from a single sample.  We
also notice that the width of CIs calculated from the single sample are in 
general about the same as the CIs found from the 200-replication sampling 
distribution, but the CIs calculated from the single sample are shifted compared
to the CIs from the sampling distribution.